*Prochlorococcus marinus* responses to light and oxygen

### 2024-03-01 Code/Manuscript_O2.Rmd
Fixed some fig captions and citations

NATLProchloroPhotosynthDepthO2_cap <- glue("**Ocean detection of *Prochlorococcus marinus* photosynthesis complexes.**", 
                       "Protein detections are plotted vs. O~2~ (µM) (X axis) and depth (m) (Y axis) at sample origin.",
                       "Rows separate data annotated as from *Prochlorococcus marinus* strains MED4 (Clade HLI), SS120 (Clade LLIII) and MIT9313 (Clade LLIV).",
                       "Columns show detections of proteins annotated as  Photosystem II (PSII), Cytochromeb6f complex (Cytb6f), Photosystem I (PSI), ATP Synthase or Ribulose-1,5-bisphosphate oxygenase carboxylase (RUBISCO).",
                       "For comparison culture growth experimental conditions are indicated by horizontal grey lines for depths approximating imposed peak Photosynthetically Active Radiation (µmol photons m^-2^ s^-1^); and vertical grey lines for [O~2~] (µM).",
                       "Point colours identify *Prochlorococcus* clade assignments",
                       "Data obtained from OceanProteinPortal (https://www.oceanproteinportal.org/).",
 .sep = " ")

@@

Guided by the evidence of ocean distributions of proteins from *Prochlorococcus* we set up a matrix of photoperiods, PAR, spectral bands, and [O~2~] to approximate current, and potential future, latitudinal, depth and seasonal niches for *Prochlorococcus* strains.  We earlier found [@murphyPhotoinactivationPhotosystemII2017] that growth under red light could prove mechanistically informative as to factors limiting *Prochlorococcus* growth, so we included that spectral band even though it is not characteristic of representative *Prochlorococcus* niches.  Although *Prochlorococcus* is currently limited to a narrow range of surface photoperiods, potential poleward latitudinal expansions, in combination with attenuation of light with depth, mean *Prochlorococcus* may potentially encounter a wide range of photoperiods.  Our growth rate determinations generally agreed with those from @mooreComparativePhysiologySynechococcus1995, for white LED and 250 µM O~2~, but our study is, to our knowledge, the first to analyze the interactive growth responses of *Prochlorococcus* strains to varying [O~2~], spectral wavebands and photoperiods.

@@

NATLProchloroProteinMetabDepthO2_cap <- glue("**Ocean detection of *Prochlorococcus marinus* protein metabolism complexes.**", 
                       "Protein detections are plotted vs. O~2~ (µM) (X axis) and depth (m) (Y axis) at sample origin.",
                       "Rows separate data annotated as from *Prochlorococcus marinus* strains MED4 (Clade HLI), SS120 (Clade LLIII) and MIT9313 (Clade LLIV).",
                       "Columns show detections of proteins annotated as FtsH Protease Complexes (FtsH1, FtsH2, FtsH3) or the Ribosome).",
                       "Culture growth experimental conditions indicated by horizontal grey lines for depths approximating Photosynthetically Active Radiation (µmol photons m^-2^ s^-1^) and vertical grey lines for [O~2~] (µM).",
                       "Point colours identify *Prochlorococcus* clade assignments",
                       "Data obtained from OceanProteinPortal (https://www.oceanproteinportal.org/).",
 .sep = " ")

@@

ProchlorococcusEnzymeKms_cap <- glue("**K~m~ values for oxygen metabolizing enzymes.**" ,
                                     "The y-axis represents the log10 concentration of oxygen substrate (mM).",
                                     "The x-axis represents the oxygen metabolizing enzymes encoded in at least one of the *Prochlorococcus marinus* strains in this study.",
                                     "The *Prochlorococcus marinus* strains are indicated in rows." ,
                                     "The solid circles represent K~m~ values from literature and the asterisks represent predicted values.",
                                     "colours represent the gene counts.",
                                     "The red shaded area denotes a K~m~ oxygen concentration range from 230 to 280 µM.",
                                     "The green shaded area denotes a K~m~ oxygen concentration range from 5 to 50 µM . ",
                                     "The blue shaded area denotes a K~m~ oxygen concentration range from 0.5 to 5 µM.  ",
                                     "The black bars show the minimum and maximum K~m~ values.",
                       "Figure was generated using a filtered subset of the annotated phytoplankton gene sequences dataset from Omar *et al*. [@omarAnnotationGenesEncoding2023].",

@@

@westNichePartitioningProchlorococcusPopulationsStratified1999 and @malmstromTemporalDynamicsProchlorococcus2010 found that decreased abundances of the LL clades corresponded to increased depth of the surface mixed layer. @malmstromTemporalDynamicsProchlorococcus2010 attributes the transport of LL ecotypes to the surface and consequent exposure to photoinhibitory high light levels as the reason for low cell abundances with increased mixed layer depth. @westNichePartitioningProchlorococcusPopulationsStratified1999 found the depth of the mixed layer strongly influenced the depth transition from HL to LL clades, but that factors other than light levels may influence the variations in the upper and lower depth limits of these ecotypes. We hypothesize that low cell abundances of LL ecotypes in the mixed layer is likely driven in part by increased [O~2~],  and it is [O~2~] that constrains LL clades to deeper waters, not necessarily the light level. We found that under 25 µM O~2~ representatives of 'LL' clades, SS120 and MIT9313, actually tolerate approximately 1.0 x 10^6^ µmol photons m^-2^ d^-1^ of PUR [@morelAvailableUsableStored1978],  comparable to the representative HL clade, MED4 which also exhibited growth saturation at the same cumulative diel PUR of 1.0 x 10^6^ µmol photons m^-2^ d^-1^.  Growth under lower O~2~ allowed MIT9313 to substantially increase its exploitation of higher diel PUR.

>>>>

NATLProchloroPhotosynthDepthO2_cap <- glue("**Ocean detection of *Prochlorococcus marinus* photosynthesis complexes.**", 
                       "Protein detections are plotted vs. O~2~ (µM) (X axis) and depth (m) (Y axis) at sample origin.",
                       "Rows separate data annotated as from *Prochlorococcus marinus* strains MED4 (Clade HLI), SS120 (Clade LLIII) and MIT9313 (Clade LLIV).",
                       "Columns show detections of proteins annotated as  Photosystem II (PSII), Cytochromeb6f complex (Cytb6f), Photosystem I (PSI), ATP Synthase or Ribulose-1,5-bisphosphate oxygenase carboxylase (RUBISCO).",
                       "For comparison culture growth experimental conditions are indicated by horizontal grey lines for depths approximating imposed peak Photosynthetically Active Radiation (µmol photons m^-2^ s^-1^); and vertical grey lines for [O~2~] (µM).",
                       "Data obtained from OceanProteinPortal (https://www.oceanproteinportal.org/).",
 .sep = " ")

@@

Guided by the evidence of ocean distributions of proteins from *Prochlorococcus* we set up a matrix of photoperiods, PAR, spectral bands, and [O~2~] to approximate current, and potential future, latitudinal, depth and seasonal niches for *Prochlorococcus* strains.  We earlier found [@murphyPhotoinactivationPhotosystemII2017] that growth under red light could prove mechanistically informative as to factors limiting *Prochlorococcus* growth, so we included that spectral band even though it is not characteristic of representative *Prochlorococcus* niches.  Although *Prochlorococcus* is currently limited to a narrow range of surface photoperiods, potential poleward latitudinal expansions, in combination with attenuation of light with depth, mean *Prochlorococcus* may potentially encounter a wide range of photoperiods.  Our growth rate determinations generally agreed with those from Moore *et al*. [@mooreComparativePhysiologySynechococcus1995], for white LED and 250 µM O~2~, but our study is, to our knowledge, the first to analyze the interactive growth responses of *Prochlorococcus* strains to varying [O~2~], spectral wavebands and photoperiods.

@@

NATLProchloroProteinMetabDepthO2_cap <- glue("**Ocean detection of *Prochlorococcus marinus* protein metabolism complexes.**", 
                       "Protein detections are plotted vs. O~2~ (µM) (X axis) and depth (m) (Y axis) at sample origin.",
                       "Rows separate data annotated as from *Prochlorococcus marinus* strains MED4 (Clade HLI), SS120 (Clade LLIII) and MIT9313 (Clade LLIV).",
                       "Columns show detections of proteins annotated as FtsH Protease Complexes (FtsH1, FtsH2, FtsH3) or the Ribosome).",
                       "Culture growth experimental conditions indicated by horizontal grey lines for depths approximating Photosynthetically Active Radiation (µmol photons m^-2^ s^-1^) and vertical grey lines for [O~2~] (µM).",
                       "Data obtained from OceanProteinPortal (https://www.oceanproteinportal.org/).",
 .sep = " ")

@@

ProchlorococcusEnzymeKms_cap <- glue("**K~m~ values for oxygen metabolizing enzymes.**" ,
                                     "The y-axis represents the log10 concentration of oxygen substrate (µM).",
                                     "The x-axis represents the oxygen metabolizing enzymes encoded in at least one of the *Prochlorococcus marinus* strains in this study.",
                                     "The *Prochlorococcus marinus* strains are indicated in rows." ,
                                     "The solid circles represent K~m~ values from literature and the asterisks represent predicted values.",
                                     "colours represent the gene counts.",
                                     "The red shaded area denotes a K~m~ oxygen concentration range from 230 to 280 µM.",
                                     "The green shaded area denotes a K~m~ oxygen concentration range from 5 to 50 µM . ",
                                     "The blue shaded area denotes a K~m~ oxygen concentration range from 0.5 to 5 µM.  ",
                                     "The black bars show the minimum and maximum K~m~ values.",
                       "Figure was generated using a filtered subset of the annotated phytoplankton gene sequences dataset from Omar *et al*. [@omarAnnotationGenesEncoding2023].",

@@

West *et al*. [@westNichePartitioningProchlorococcusPopulationsStratified1999] and Malmstrom *et al*. [@malmstromTemporalDynamicsProchlorococcus2010] found that decreased abundances of the LL clades corresponded to increased depth of the surface mixed layer. Malmstrom *et al*. [@malmstromTemporalDynamicsProchlorococcus2010] attributes the transport of LL ecotypes to the surface and consequent exposure to photoinhibitory high light levels as the reason for low cell abundances with increased mixed layer depth. @westNichePartitioningProchlorococcusPopulationsStratified1999 found the depth of the mixed layer strongly influenced the depth transition from HL to LL clades, but that factors other than light levels may influence the variations in the upper and lower depth limits of these ecotypes. We hypothesize that low cell abundances of LL ecotypes in the mixed layer is likely driven in part by increased [O~2~],  and it is [O~2~] that constrains LL clades to deeper waters, not necessarily the light level. We found that under 25 µM O~2~ representatives of 'LL' clades, SS120 and MIT9313, actually tolerate approximately 1.0 x 10^6^ µmol photons m^-2^ d^-1^ of PUR [@morelAvailableUsableStored1978],  comparable to the representative HL clade, MED4 which also exhibited growth saturation at the same cumulative diel PUR of 1.0 x 10^6^ µmol photons m^-2^ d^-1^.  Growth under lower O~2~ allowed MIT9313 to substantially increase its exploitation of higher diel PUR.
